---
title: 'Homework 4: Bags, Forests, Boosts, oh my'
author: "Will Hall"
date: "2/28/2019"
output: github_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(glmnet)
library(caret)
library(ISLR)
library(janitor)
library(stringr)
library(rpart)
library(rpart.plot)
library(partykit)
library(randomForest)
library(MASS)
library(gbm)
theme_set(theme_bw())

```

## Problem 1

Problem 7 from Chapter 8 in the text. To be specific, please use a sequence of
`ntree` from 25 to 500 in steps of 25 and `mtry` from 3 to 9 for by 1. 

## Answer 1

#Conclusive Random Forest with Corresponding MSEs:

```{r}
set.seed(690705)
df <- tbl_df(Boston)

for (k in 1:20){
  inTraining <- createDataPartition(df$medv, p = .75, list = F)
  training <- df[inTraining, ]
  validating <- df[-inTraining, ]}
  
  mtry <- seq(3,9, by = 1)
  ntree <- seq(25, 500, by = 25)
  
  conclsive_rf <- list(mtry = mtry, ntree = ntree) %>% cross_df()
  
  get_mse <- function(mtry, ntree){
      rf_train <- randomForest(medv ~ .,
                               data = training,
                               mtry = mtry,
                               ntree = ntree)
      mse <- mean((predict(rf_train, newdata = validating) - validating$medv)^2)
      tibble(
        mtry = mtry,
        ntree = ntree,
        mse = mse
      )
  }
  
  
  c_rf_mse <- pmap_df(conclsive_rf, get_mse)
  library(magrittr)
  c_rf_mse %<>% arrange(mtry)
  View(c_rf_mse)
  
```

#Graphical Display: 
  
```{r}  
c_rf_mse %>% ggplot(aes(x=ntree, y=mse, group=mtry, color=factor(mtry))) +geom_line() 
  
```
  
  
#Model Interpretation: 

Through this model, 5 appears to be the best random predictor variable limit for random forests. The display shows 5’s corresponding MSE values stay consistently lower relative to other limits. And from the lab, even though higher limits display lower MSEs through cross-validation, it is my intention to avoid overfitting. So after interpreting the model above, and keeping in mind the cross-validation run, 5 is the ideal mtry.



#Fitting the model with the best random predictor variable (5):

```{r}
rf_boston_5 <- randomForest(medv ~ ., 
                            data = training,
                            mtry = 5)
rf_boston_5


test_preds <- predict(rf_boston_5, newdata = validating)
boston_test_df <- validating %>%
  mutate(y_hat_rf_5 = test_preds,
         sq_err_rf_5 = (y_hat_rf_5 - medv)^2)

```

#Test MSE for 5:

```{r}
mean(boston_test_df$sq_err_rf_5)
```

#Understanding variable importance through the best random predictor variable (5):

```{r}

rf_boston_5_grid <- train(medv ~ ., 
                      data = training,
                      method = "rf",
                      ntree = 500,
                      importance = T,
                      tuneGrid = data.frame(mtry = 3:9))
rf_boston_5_grid

imp <- varImp(rf_boston_5_grid)$importance
rn <- row.names(imp)
imp_df <- data_frame(variable = rn, 
                     importance = imp$Overall) %>%
  arrange(desc(-importance)) %>%
  mutate(variable = factor(variable, variable))

p <- ggplot(data = imp_df,
            aes(variable, importance))
p + geom_col(fill = "#6e0000") +
  coord_flip()
```

=======
## Answer 1

>>>>>>> 0fb048fc23b65954eedcba57aa4bcd07b5c80252
## Problem 2

Problem 8 from Chapter 8 in the text. Set your seed with 9823 and split into 
train/test using 50\% of your data in each split. In addition to 
parts (a) - (e), do the following:

1. Fit a gradient-boosted tree to the training data and report the estimated 
test MSE. 
2. Fit a multiple regression model to the training data and report the 
estimated test MSE
3. Summarize your results. 
<<<<<<< HEAD

## Answer 2

#Loading Data:

```{r}
df <- tbl_df(Carseats) %>%  na.omit()
```

#Splitting Training and Validating Set:

```{r}
set.seed(9823)
inTraining <- createDataPartition(df$Sales, p = .5, list = F)
training <- df[inTraining, ]
validating  <- df[-inTraining, ]

```

#Fitting a Regression Tree to the Validation Data:

```{r}
tree_carseats <- rpart(Sales ~ ., data = training)

```

#Summary and Plot

```{r}
summary(tree_carseats)
plot(as.party(tree_carseats))

```

#MSE for Tree:

```{r}
pred_carseats <- predict(tree_carseats, validating)
mean((validating$Sales - pred_carseats)^2)
```

#Cross-validation for Carseats Using Random Forest

```{r}
rf_carseats_cv <- train(Sales ~ ., 
                      data = training,
                      method = "rf",
                      ntree = 125,
                      importance = T,
                      tuneGrid = data.frame(mtry = 1:10))
rf_carseats_cv
```

#Plotted Random Forest CV

```{r}
plot(rf_carseats_cv)

```

#Test MSE for Random Forest: mtry = 4

```{r}
#I chose 4 based on 5's small relative gain in R2, and small relative drop in RMSE to 4. Using mtry = 4 within random forest, compared to the initial tree regression, ends up lowering the test MSE from 4.48 to 3.28.

rf_carseats_4 <- randomForest(Sales ~ ., 
                          data = training,
                          mtry = 4,
                          importance = T)
                    
rf_carseats_4



test_preds <- predict(rf_carseats_4, newdata = validating)
carseats_test_df <- validating %>%
  mutate(y_hat_rf_4 = test_preds,
         sq_err_rf_4 = (y_hat_rf_4 - Sales)^2)


pred_carseats4 <- predict(rf_carseats_4, validating)
mean((validating$Sales - pred_carseats4)^2)


```

#Variable Importance:

```{r}
importance(rf_carseats_4)

#Here, ShelveLoc and Price are also the most important variables
```

#Graphical Representation of Variables via Best rf mtry (4):

```{r}
set.seed(1982)

rf_carseats_4_grid <- train(Sales ~ ., 
                      data = training,
                      method = "rf",
                      ntree = 500,
                      importance = T,
                      tuneGrid = data.frame(mtry = 1:4))


imp <- varImp(rf_carseats_4_grid)$importance
rn <- row.names(imp)
imp_df <- data_frame(variable = rn, 
                     importance = imp$Overall) %>%
  arrange(desc(-importance)) %>%
  mutate(variable = factor(variable, variable))
p <- ggplot(data = imp_df,
            aes(variable, importance))
p + geom_col(fill = "#6e0000") +
  coord_flip()
```

#Bagging Approach:

```{r}
set.seed(9823)
bagged_carseats <- randomForest(Sales ~ ., data = training, mtry = 10)
bagged_carseats
```

#Estimating the Bagged MSE:

```{r}

test_preds <- predict(bagged_carseats, newdata = validating)
bagged_carseats_test_df <- validating %>%
  mutate(y_hat_bags = test_preds,
         sq_err_bags = (y_hat_bags - Sales)^2)

mean(bagged_carseats_test_df$sq_err_bags)
```

#Importance of Bagged Variables:

```{r}
set.seed(9823)
bagged_variables <- randomForest(Sales ~ ., data = training, mtry = 10, importance = T)
importance(bagged_variables)

#ShelveLoc and Price are the most important variables
```

#Gradient-boosting Approach:

```{r}
set.seed(1982)
grid <- expand.grid(interaction.depth = c(1, 3), 
                    n.trees = seq(0, 2000, by = 100),
                    shrinkage = c(.01, 0.001),
                    n.minobsinnode = 10)
trainControl <- trainControl(method = "cv", number = 6)
gbm_carseats <- train(Sales ~ ., 
                    data = training, 
                    distribution = "gaussian", 
                    method = "gbm",
                    trControl = trainControl, 
                    tuneGrid = grid,
                    verbose = FALSE)
gbm_carseats
```

#Graphical Representation of Gradient-boosted Tree (mtry = 6):

```{r}
plot(gbm_carseats)
```

#MSE for Gradient-boosted Tree:

```{r}

test_preds <- predict(gbm_carseats, newdata = validating)
carseats_test_df <- carseats_test_df %>%
  mutate(y_hat_gbm = test_preds,
         sq_err_gbm = (y_hat_gbm - Sales)^2)

mean(carseats_test_df$sq_err_gbm)
```

#Training a Multiple Regression (OLS):

```{r}
lm_carseats <- lm(Sales ~ ., data = training)
```

#MSE for OLS Model:

```{r}

test_preds <- predict(lm_carseats, newdata = validating)
carseats_test_df <- carseats_test_df %>%
  mutate(y_hat_gbm = test_preds,
         sq_err_gbm = (y_hat_gbm - Sales)^2)

mean(carseats_test_df$sq_err_gbm)
```

#Gradient-boosted Tree vs. OLS Model MSE Results:

The OLS Model actually has a lower MSE than the Gradient-boosted Tree model. This result demonstrates that more simple models should not be discounted in lieu of so many other complex instruments. Context and simple model comparisons must also play a key role in  choice. 
